Appendix

### 2020-05-26 manuscript/bef_stability/appendix.Rmd
Fix PCA plot

---
title: Appendix of Complex mecanisms linking biomass stability and total biomass with species richness and network structure in stream ecosystems 
author: Alain Danet, Maud Mouchet, Elisa Thebault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{float}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

```{r, echo = FALSE, message = FALSE, collapse = FALSE}
mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE
)
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(7, 7),
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "total_sem_effect.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))

@@

```{r}
# Load manuscript results
load(file = mypath("manuscript", "bef_stability", "result", "workspace.rda"))
```

@@

```{r}

The fish community sampling took place in a variety of stream size, shape and
depth. This diversity constraints the type of protocol that can take place.
For the small streams (depth < 0.7m across the entire width), the complete protocol
is used. It consists in placing anodes every X meter across the width of the
river. **Add the precise procedure?**
The fishermen made from one to three row of fishing. We kept only the first row
for analysis.
For deeper and wider streams, the sampling was partial over the bank and the
channel and performed by boat and by foot (along the bank). We kept only
sampling done over the bank because channel sampling was not always done.
Two types of partial protocols takes place: partial over bank and partial by
points. The former is continuous while the latter is discrete. Furthermore, in
partial by point protocol the channel can be sampled as well. We kept only the
partial by point fish sampling which all the sampling point were done over the
bank to homogenize the sampling. Then, we merged partial over bank and partial
over point protocols. Some sampling stations had mix characteristics in which mix of
partial and complete protocols were applied depending of the depth of the river
at the time of sampling. Those stations were removed from the analysis.

### Batches definition and body size measurement 
During the electric fishing, and before that fish body size are measured, they
are regrouped by batches. Each batches contains one species but there can be
several batches by species. Body size assessment are different according to the
batches. There are four types of batches: N, I, S/L and G. The batches N and
I contains the biggest individuals, there are normally less numerous and then
the body size of the individuals of those batches is recorded. The batches S/L
contains fishes of intermediate sizes. Inside each S/L batches, the operator
tries to make batches as homogeneous as possible in body size. Then, the body
size of a subsample of X individuals are recorded. The G batches contain
individuals of low body size, with care that body size is as homogenous as
possible in the batches. Then body size of the shortest and the longest
individuals are recorded.

@@

fish_length %<>%
  left_join(lot_id_opcod) %>%
  rename(length = fish) %>%
  select(opcod, species, length)
length_sem_analysis <- fish_length %>% 
  filter(opcod %in% op_analysis_bbb$opcod)
na_length <- sum(is.na(length_sem_analysis$length))
```

@@

We removed batches that did not satisfied the following conditions. First, we
removed batches which the total number of individuals was not recorded. In lot
"G", we removed batches were  $l_{min} \geq l_{max}$ or if at least one of the
measurement was missing. Additionally, G batches were removed when number of
individuals fell below five, as G batches normally contained more individuals.
In the S/L lot, we removed batches were the number of individuals in the
recorded individuals fell below 10. In sum, those sanity check allowed to
exclude potentially problematic batches. Finally, in the dataset used in the
analysis, `r na_length` individuals had missing length over 
`r nrow(fish_length)` individuals.

@@

## Network inference

@@

@@

** I question myself if I should move the network inference figure here and
just let the temporal network in the main text. May be even add the temporal
biomass plot (species/trophic species) in the main text**

@@

```{r}
myload(fourier, dir = mypath("data-raw", "polluants"))
# Where is DBO?
```

For BOD, we removed measurements that were expressed in uncommon units.

The first goal was to harmonize the different variables and scale it to the
timescale of fish sampling, we took annual average of each variables. To reduce 
daily or seasonal effects, we first applied average data with the moving
window of one year. For temperature, the moving was performed on daily average
temperature. Then we took the annual average of each year and each variable.

@@

Then we modeled each variable as dependant only on the covariation between sites
and taking in account stream network. To model the autocovariance between sites,
we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their performance of prediction. The interpolations
were done with the `SSN` R package 
[@isaak_applications_2014, @hoef_ssn_2014].

@@

```{r eval = FALSE}
myload(cv_press_interp_mv_avg,
  dir = mypath("data-raw", "polluants"))
myload(cv_temp_press_interp_mv_avg,
  dir = mypath("data-raw", "naiades_temperatures"))
myload(cv_flow_press_interp_mv_avg,
  dir = mypath("data-raw", "flow"))
dbo_cv <- cv_press_interp_mv_avg %>%
  filter(parameter == "dbo5")

cv_dbo_temp_flow <- map2_dfr(
  list(dbo = dbo_cv, temp = cv_temp_press_interp_mv_avg, flow = cv_flow_press_interp_mv_avg),
  c("BOD", "Temperature", "Flow"),
  function(x, y) {
    x %>%
      group_by(basin, year) %>%
      summarise_if(is.numeric, list(avg = ~mean(., na.rm = TRUE))) %>%
      mutate(Variable = y) %>%
      select(Variable, basin, year, everything()) %>%
      ungroup
  })

@@

```{r, fig.show = "hide"}
axis_pair <- list(c(1,2), c(1,3), c(2,3), c(1, 4), c(2, 4), c(3,4), c(4,5))
p_l_rot <- map(axis_pair, 
  function (axis)
    pca_rotated_plot <- plot_rotated_pca(
      pca_rotated = pca_rotated,
      axis = axis 
    )
)
```

@@

pca_caption <- paste0(
  "Rotated axis of PCA on environmental variables. The five components
  explained,", tot_var_expl_pca, "% of the variance (",
  paste0(names(prop_var_expl_pca), ":", prop_var_expl_pca, "%"), ").",
  "(A) Axis 1 was positively related to the river size, axis 2 was related to
  average temperature and negatively related to the river slope and altitude.
  (B) Axis 3 was negatively related to the CV of flow and average enrichment.
  Axis 4 was related to the CV of river size. Finally, axis 5 was related to the
  CV of enrichment.")
```

```{r pca-rotated, fig.cap = pca_caption}
to_plot <- c(1, 6, 7)
plot_grid(plotlist = map(p_l_rot[to_plot], ~.x$rotated),
  ncol = length(to_plot), labels = LETTERS[1:3])
```

@@

@@

####  SEM indirect and total effect

### Model linear Relationship

```{r}
hist_var_caption <- paste0(
  "Distribution of the variables used in the Structural Equation Modeling."
)
```

```{r hist-sem-data, fig.cap = hist_var_caption}
sem_data %>%
  select(log_bm_std, log_rich_tot_std, log_sync, log_stab_std, log_cv_sp, log_RC1, log_RC2, log_RC3, RC4,
    RC5, t_lvl, ct, log_bm) %>%
  gather(variable, value) %>%
  mutate(variable = str_replace_all(variable, get_sem_var_name_replacement())) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free")
```

@@

kable(colin_stab_sem,
  format = "latex",
  booktabs = T,
  caption = caption_indir_bm_sem
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

@@

```{r}

est_bm_sem_caption <- paste0(
  "Estimate of the effects for the total biomass structural equation
  modeling."
)

@@

```{r tab-coeff-bm, fig.cap = "Coefficient"}

# Replacement rules for term
rp_est <- c(
  "(Intercept)" =     "intercept",
  "sd\\_\\(intercept\\)\\.basin" = "sd intercept basin",
  "log_rich" =     "Median richness (log base 10)",
  "troph_group3" =     "high trophic group",
  "log_rich:troph_group3" =     "Median richness (log10) : high trophic group",
  "sig01.basin" =     "sd intercept basin",
  "sd_Observation.Residual" =     "residuals",
  "sigma" = "residuals"
)
# Replacement rules for group
rp_group <- c(
  "fixed" = "Fixed",
  "basin" = "Random"
)

@@

>>>>

---
title: Appendix
author: Alain Danet, Maud Mouchet, Elisa Thebault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{float}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

```{r, echo = FALSE, message = FALSE, collapse = FALSE}

mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

load(file = mypath("manuscript", "bef_stability", "result", "workspace.rda"))

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE 
)

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "total_sem_effect.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))

@@

```{r}
# Load manuscript results
```

@@

```{r}
```

The fish community sampling took place in a variety of stream size, shape and
depth. This diversity constraints the type of protocol that can take place.  For
the small streams (depth < 0.7m across the entire width), the complete protocol
is used. It consists in placing anodes every X meter across the width of the
river (**Add the precise procedure?**). The fishermen made from one to three row
of fishing. We kept only the first row for analysis to homogenize the sampling
effort. For deeper and wider streams, the sampling was partial over the bank and
the channel and performed by boat and by foot (along the bank). We kept only
sampling done over the bank because channel sampling was not always done.  Two
types of partial protocols takes place: partial over bank and partial by points.
The former is continuous while the latter is discrete. Furthermore, in partial
by point protocol the channel can be sampled as well. We kept only the partial
by point fish sampling which all the sampling point were done over the bank to
homogenize the sampling.  Then, we merged partial over bank and partial over
point protocols. Some sampling stations had mix characteristics in which mix of
partial and complete protocols were applied depending of the depth of the river
at the time of sampling. Those stations were removed from the analysis.

Heterogeneity in the sampling period may introduce variability that is not related
to ecological processes such as the presence of juveniles. In order to reduce
heterogeneity in sampling time, only one sampling per year was retained if there
was two in a given year, the one with with the most species and then individual
collected. Finally, most of the fish sampling occurred in September (Median:
mid-August) and the median standard deviation of the sampling month was inferior
to one (0.7) within station. Additionally, we removed sampling events, in a
given station, which the sampled length along the stream was outside the median
length sampled more or less 30%. Finally, we standardadized the biomass record
and species richness by the sampled surface to get rid of sampling effort
effects.

### Batches definition and body size measurement

During the electric fishing, and before that fish body size are measured, they
are regrouped by batches. Each batches contains one species but there can be
several batches by species. Body size assessment are different according to the
batches. There are four types of batches: N, I, S/L and G. The batches N and
I contains the biggest individuals, there are normally less numerous and then
the body size of the individuals of those batches is recorded. The batches S/L
contains fishes of intermediate sizes. Inside each S/L batches, the operator
tries to make batches as homogeneous as possible in body size. Then, the body
size of a subsample of X individuals are recorded. The G batches contain
individuals of low body size, with care that body size is as homogenous as
possible in the batches. Then body size of the shortest and the longest
individuals are recorded.

@@

fish_length %<>%
  left_join(lot_id_opcod) %>%
  rename(length = fish) %>%
  select(opcod, species, length)

length_sem_analysis <- fish_length %>% 
  filter(opcod %in% op_analysis_bbb$opcod)
rm(fish_length, lot_id_opcod)
nrow_fish_length <- nrow(length_sem_analysis)
na_length <- sum(is.na(length_sem_analysis$length))
```

@@

We removed batches that did not satisfied the following conditions. First, we
removed batches which the total number of individuals was not recorded. In lot
"G", we removed batches were  $l_{min} \geq l_{max}$ or if at least one of the
measurement was missing. Additionally, G batches were removed when number of
individuals fell below five, as G batches normally contained more individuals.
In the S/L lot, we removed batches were the number of individuals in the
recorded individuals fell below 10. In sum, those sanity check allowed to
exclude potentially problematic batches. Finally, in the dataset used in the
analysis, `r na_length` individuals had missing length over 
`r nrow_fish_length` individuals.

@@

### Metaweb

#### Ontogenic diet

Diet information were obtained from fishbase.org and literature [REF willem?]. The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had as strict piscivorous
life stage. Fish species were largely omnivorous.

#### Define trophic interactions

```{r}
trophic_building_caption <- paste0(
  "Trophic species"
) 
```

```{r p-net-inf, fig.cap=trophic_building_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "network_inference.pdf"))
```

@@

The metaweb described the potential trophic interactions between the
($`r ncol(metaweb_analysis$metaweb) - length(metaweb_analysis$resource)`$) trophic species.
We had seven resource nodes which were present in the diet of the trophic
species namely detritus, biofilm, phytoplankton, zooplankton, macrophages,
phytobenthos and zoobenthos.

@@

A trophic species belonged to a life stage if the median of its body size range
fell in the range of the life stage. In turn, this trophic species has feeding
interactions with the resource nodes of its life stage. Interactions between
resource nodes were determined by literature. For example, phytobenthos fed on
detritus, while zooplankton fed on zooplankton and phytoplankton. Fish-fish
trophic interactions between a predator and a prey were determined by body size
ratio as in many studies [@brose_predator_2019; @poisot_structure_2016;
@schneider_body_2012].  For a given trophic species, we first determined if its
size class end up in a life stage containing piscivorous diet, i.e. if there was
an overlap between them. We were more liberal for piscivory determination to not
miss potential fish-fish trophic interactions which were the main interactions
in our community. If the trophic species was piscivorous, we determined its prey
size range from 3% to 45% of the median of its size range. A trophic interaction
was set for each trophic species whose median of its size class fell into the
prey range of the predator trophic species. To sum up, the metaweb was obtained
with allometric feeding interactions (1) between trophic species, (2) between
trophic species and resource nodes and (3) between resource nodes.

#### Obtain network of a community

The trophic network of a community (i.e. one network for each fish community
sampling) was obtained by determining the trophic species of each fish
individual in the community and interactions were set according to the metaweb.
All the resource nodes were assumed to be present in each local network.
Finally, the structure of the local networks varied according the presence of
the trophic species. The inferred interactions represent potential interactions
(i.e. not measured interactions) and are binary (i.e. presence or absence of
feeding interactions).

@@

```{r, fig.height = 20/2.54}
myload(fourier, dir = mypath("data-raw", "polluants"))
myload(flow_temp_metrics, dir = mypath("data"))
sf_pressure <- station_analysis %>% 
  select(id) %>%
  left_join(rename(pca_pressure, id = station)) %>%
  na.omit()

p <- ggplot() +
  geom_sf(data = the_8_hydrologic_basin)

var_to_plot <- colnames(sf_pressure)[
  !colnames(sf_pressure) %in% c("id", "geometry")]
var_replacement <- c(
  "slope" = "Slope",
  "alt" = "Altitude",
  "d_source" = "Distance from source",
  "strahler" = "Strahler order",
  "width_river_mean" = "Average stream width",
  "avg_depth_station_mean" = "Average stream depth",
  "width_river_cv" = "CV of stream width",
  "avg_depth_station_cv" = "CV of stream depth",
  "DBO_med" = "Average BOD",
  "flow_med" = "Average flow",
  "temperature_med" = "Average temperature",
  "DBO_cv" = "CV of BOD",
  "flow_cv" = "CV of flow",
  "temperature_cv" = "CV of temperature"
)

map_pressure <- map(var_to_plot,
~p +
  geom_sf(data = sf_pressure, aes_string(color = .x)) +
  labs(title = str_replace_all(.x, var_replacement)) +
  theme(legend.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
)

plot_grid(plotlist = map_pressure, ncol = 2)

The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD resp.). For each recording
station, we applied a moving window averaging (more details in supplementary
material) to get rid of seasonal effects. For temperature, the moving was performed on daily
average temperature. Then we took the annual average of each year and each
variable (i.e. one value by year and by station).

@@

Then we modeled each variable as dependant only on the covariation between sites
and taking in account stream network. To model the autocovariance between sites,
we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their prediction performance. The interpolations
were done with the `SSN` R package 
[@isaak_applications_2014, @hoef_ssn_2014].

@@

```{r eval = TRUE}
myload(cv_press_interp_mv_avg,
  dir = mypath("data-raw", "polluants"))
myload(cv_temp_press_interp_mv_avg,
  dir = mypath("data-raw", "naiades_temperatures"))
myload(cv_flow_press_interp_mv_avg,
  dir = mypath("data-raw", "flow"))
dbo_cv <- cv_press_interp_mv_avg %>%
  filter(parameter == "dbo5")

cv_dbo_temp_flow <- map2_dfr(
  list(dbo = dbo_cv, temp = cv_temp_press_interp_mv_avg, flow = cv_flow_press_interp_mv_avg),
  c("BOD", "Temperature", "Flow"),
  function(x, y) {
    x %>%
      select(basin, year, bias, std.bias, RMSPE, std.MSPE) %>%
      group_by(basin) %>%
      summarise_if(is.numeric, list(avg = ~mean(., na.rm = TRUE), sd = ~sd(., na.rm = TRUE))) %>%
      mutate(Variable = y) %>%
      select(Variable, basin, everything()) %>%
      ungroup
  })

@@

```{r eig-plot}
rotated_eigen_values <- 
  tibble(
    name = paste0("RC", seq(1,14)),
    type = c(rep("Kept", 5), rep("No kept", 9)),
    value = pca_rotated$rotated$values
  )
eigen_rotated_plot <-
  ggplot(rotated_eigen_values,
    aes(x = reorder(name, -value) , y = value)) +
  geom_col(aes(fill = type)) +
  scale_fill_discrete(name = "Status") +
  labs(x = "Rotated component", y = "Eigen values")
```

```{r contrib-pca}
#https://stats.stackexchange.com/a/143949
caption_contrib_pca <- paste0(
  "Loadings of variable to rotated components of the PCA.
  Loadings are the coefficients which determine the components of the PCA as
  linear combinations of the variables.
  "
)

loadings_pca <- matrix(
  as.matrix(round(pca_rotated$rotated$Structure, 2)),
  nrow = 14,
  dimnames = dimnames(pca_rotated$rotated$Structure)
  ) %>%
  as_tibble %>%
  mutate(
    Variable = str_replace_all(
      dimnames(pca_rotated$rotated$Structure)[[1]],
      pca_var_replacement)
  ) %>%
  select(Variable, RC1, RC2, RC3, RC4, RC5)

kable(
  x = loadings_pca,
  format = "latex",
  booktabs = T,
  label = "contrib-pca",
  caption = caption_contrib_pca
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r}
corr_pca_caption <- paste0(
  "Spearman correlation between environmental variables used in the PCA.
  The variables have been scaled before to compute the correlation."
)

pca_var_correlation <- pca_pressure %>%
  select(-station) %>%
  #mutate_all(scale) %>%
  as.matrix() %>%
  cor(., method = "spearman")

pca_var_correlation %<>% round(., 2)
dimnames(pca_var_correlation) %<>%
  map(., ~str_replace_all(.x, get_pca_var_name_replacement()))

kable(
  x = pca_var_correlation,
  format = "latex",
  booktabs = T,
  label = "corr-pca",
  caption = corr_pca_caption
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

@@

pca_caption <- paste0(
  "Rotated axis of PCA on environmental variables. The five components
  explained,", tot_var_expl_pca, "% of the variance (",
  paste0(names(prop_var_expl_pca), ":", prop_var_expl_pca, "%"), ").",
  "(A) Axis 1 was positively related to the river size, axis 2 was related to
  average temperature and negatively related to the river slope and altitude.
  (B) Axis 3 was negatively related to the CV of stream size.
  Axis 4 was related to the CV of enrichment.
  Finally, axis 5 was related to the CV of flow and average enrichment.")
```

```{r pca-rotated, fig.cap = pca_caption, fig.asp = 1}
p1 <- my_pca_plot(xaxis = "RC1", yaxis = "RC2", ctb_thld = .4)
p2 <- my_pca_plot(xaxis = "RC3", yaxis = "RC4", ctb_thld = .4)
p3 <- my_pca_plot(xaxis = "RC4", yaxis = "RC5", ctb_thld = .4)

pca_plot <- plot_grid(p1, p2, p3, eigen_rotated_plot, nrow = 2)
save_plot(
  filename = mypath("manuscript", "bef_stability", "figs", "pca_plot.pdf"),
  pca_plot,
  nrow = 2, ncol = 2,
  base_height = 6/2.54, base_asp = 1, base_width = NULL 
  )
pca_plot
```

@@

```{r }
rc2 <- tibble(
  station = pca_pressure$station,
  alt = pca_pressure$alt,
  temperature_med = pca_pressure$temperature_med,
  value = pca_rotated$rotated$scores[,"RC2"]
)

rc2_sf <- station_analysis %>%
  select(station = id) %>%
  filter(station %in% rc2$station) %>%
  left_join(rc2, by = "station")

caption_rc2_map <- paste0(
  "Map of the station values for the second axis of the rotated PCA."
)
```

```{r rc2-map, fig.cap = caption_rc2_map}
p +
  geom_sf(data = rc2_sf, aes(color = log(value + abs(min(value)) + 1 ))) +
  labs(title = "RC2 (Log)") +
  theme(legend.title = element_blank())
```

```{r pca-two-axis, eval = FALSE}
two_axis <- compute_rotated_pca(.data = pca_pressure, naxis = 2)
two_axis$rotated$Vaccounted

cor(two_axis$rotated$score[,c("RC1", "RC2")],
  pca_rotated$rotated$score[,c("RC1", "RC2")])
```

@@

### Data transformation

Additionally, we logged the axis one, two and three to normalize their
distribution and we reversed the third axis to get increasing average
temperature with increasing values of the axis and so decreasing values of
altitude.

```{r}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 5)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 5), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c(rep("Log base 10", 3), rep("No", 2), "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  interpretation = c(rep("", 11)) 
) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "Variable description and transformation"
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

### SEM hypothesis

We linked environmental variables both to community structure and stability
components. Rooted in assembly theory, environment filters species that can
establish in a community and thus affect community structure. (+ add specific
hypothesis about physical and climate). (1) We hypothesized that enrichment
(characterized by DBO) and mean annual temperature will increase species turnover
[@friedpetersen_drivers_2020]. (2) We hypothesized environmental forcing can also
decrease biomass stability directly without going by community structure. First,
instability of habitat and environmental factors may in turn drive biomass
instability [@hansen_climate_2013]. Secondly, average environmental conditions
such as enrichment can increase directly CV by increasing species biomass
fluctuations [@tabi_warming_2019]. Increasing temperature was also hypothesized
to decrease stability through starvation resulting from high metabolic activity
[@tabi_warming_2019].

---
# Species richness to community attributes 
---
To reveal the mechanisms behind the relationship between species richness and
biomass stability, causal links were set from species richness to other
community attributes. In particular, species richness decreased connectance
following results from @thebault_stability_2010; @winemiller_must_1989. But
theoretically, some ecological networks can show the inverse relationship
[@winemiller_must_1989]. (6) We hypothesized also that species richness will
modulate average weighted trophic level because of food web assembly processes.
A community with more fish species can support higher trophic level. Finally, we
set up causal relationships from community attributes to stability components.

---
#  Community attributes to stability
---
We hypothesized that (7) species richness decreased synchrony between species
[@loreau_species_2008] and increased $\overline{CV_{sp}}$
[@tilman_biodiversity_2006]. (8a) We expected that average trophic level had a
positive or negative relationship with synchrony, bottom-up effects should
result in positive whereas top-down in negative relationships. We expected a (8b)
negative relationship between average trophic level and $\overline{CV_{sp}}$ if
predators stabilised prey fluctuations or if the increasing part of
biomass in the predators decreased $\overline{CV_{sp}}$ because of their higher
body mass (if mean increase faster than variance). (9a) Connectance was expected
to decrease or increase $\overline{CV_{sp}}$ and (9b) synchrony because of top
down effects or bottom up effects.

## Model linear Relationship

### Variables

```{r}
hist_var_caption <- paste0(
  "Distribution of the variables (before transformation) used in the Structural Equation Modeling."
)
```
```{r hist-sem-data, fig.cap = hist_var_caption}
sem_data %>%
  select(log_bm_std, log_rich_tot_std, log_sync, log_stab_std, log_cv_sp, log_RC1, log_RC2, log_RC3, RC4,
    RC5, t_lvl, ct, log_bm) %>%
  gather(variable, value) %>%
  mutate(variable = str_replace_all(variable, get_sem_var_name_replacement())) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free")
```

@@

kable(colin_stab_sem,
  format = "latex",
  booktabs = T,
  caption = "Variance Inflation Factor for the linear models used in structural
  equation modeling"
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

@@

```{r}
est_bm_sem_caption <- paste0(
  "Estimate of the effects for the total biomass structural equation
  modeling."
)

@@

```{r tab-coeff-bm, fig.cap = "Coefficient"}
# Replacement rules for term
rp_est <- c(
  "(Intercept)" =     "intercept",
  "sd\\_\\(intercept\\)\\.basin" = "sd intercept basin",
  "log_rich" =     "Median richness (log base 10)",
  "troph_group3" =     "high trophic group",
  "log_rich:troph_group3" =     "Median richness (log10) : high trophic group",
  "sig01.basin" =     "sd intercept basin",
  "sd_Observation.Residual" =     "residuals",
  "sigma" = "residuals"
)
# Replacement rules for group
rp_group <- c(
  "fixed" = "Fixed",
  "basin" = "Random"
)
```

@@

### Robustness to different dataset selection

```{r sensitivity, fig.dim = c(10, 7), fig.cap = "Sensivity analysis with different dataset", cache = TRUE}
library(furrr)

myload(op_analysis, op_analysis_wo_holes, op_analysis_ov, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)

#myload(network_metrics, dir = mypath("data", "classes"))
sem_sensivity_data <- list(full = op_analysis, bbb = op_analysis_bbb,
  without_holes = op_analysis_wo_holes, bbb_without_holes = op_analysis_wo_holes_bbb)

sem_stab_l_data <- furrr::future_map(sem_sensivity_data,

  ~build_dataset_get_sem_coefficient(
    .op = .x,
    sem_fun = compute_stab_sem_rich
)
  )

nb_station <- map_int(sem_sensivity_data, ~length(unique(.x$station)))

sem_stab_results <- map(sem_stab_l_data,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

# Reformat
colnames(sem_stab_results) <- str_to_lower(colnames(sem_stab_results))
sem_stab_results %>%
  mutate(r_p = str_c(response, predictor, sep = "~")) %>%
  ggplot(aes(x = r_p, y = std.estimate)) +
  geom_point(aes(color = dataset)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r sem-sens-bm, cache = TRUE, fig.cap = "Sensivity analysis for Biomass"}
sem_bm_l_data <- furrr::future_map(sem_sensivity_data,
  ~build_dataset_get_sem_coefficient(
    .op = .x,
    sem_fun = compute_prod_sem_rich)
)
sem_bm_results <- map(sem_bm_l_data,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

# Reformat
colnames(sem_bm_results) <- str_to_lower(colnames(sem_bm_results))
sem_bm_results %>%
  mutate(r_p = str_c(response, predictor, sep = "~")) %>%
  ggplot(aes(x = r_p, y = std.estimate)) +
  geom_point(aes(color = dataset)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
